---
title: "15 Years of UFOs: Reported cases of UFOs in continental US between 2000 and 2014"
subtitle: A `tidytuesday` experiment.
author: 
  - name: Tomer Zipori
    edu: BA
degree: Master of Visualization
institute: BGU
faculty: Nerds
department: Psychology
location: Be'er Sheva, IL 
date:
  - year: 2023
    month: May
sign_page: true
dedication: To my parents.
acknowledge: I acknowledge the tribal chief.
committee:
  - name: Jane Doe III
    prefix: Dr
    position: Chair of Department
  - name: John Smith
    prefix: Dr
    position: Faculty Advisor
lof: false
lot: false
toc-title: Contents
output:
  pagedown::thesis_paged: 
    toc: true
    number_sections: yes
    pandoc_args: --mathjax
    self_contained: no
link-citations: yes
# uncomment this line to produce HTML and PDF in RStudio:
knit: pagedown::chrome_print
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, warning = F,
                      fig.align = 'center')
```

# Setup {.chapter}
## Setup
```{r}
library(tidyverse)
library(readxl)
library(tidytuesdayR)
library(maps)
library(lubridate)
library(jpeg)
library(ggimage)
library(showtext)
library(flextable)
library(glue)
```


## Loading a dataset
```{r loading data}
ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")

usa <- map_data("state")

state_codes <- read_csv("data/state_code.csv") %>%
  select(state, code) %>%
  mutate(state = tolower(state), code = tolower(code))

uspop <- read_excel("data/uspop.xlsx", col_names = c("region", "pop_2010", "pop_2011", "pop_2012", "pop_2013", "pop_2014")) %>%
  mutate(region = tolower(str_remove(region, "."))) %>%
  rowwise() %>%
  mutate(mean_pop = mean(c(pop_2010, pop_2011, pop_2012, pop_2013, pop_2014))) %>%
  ungroup() %>%
  select(region, mean_pop)
```

## Functions
```{r}
convert_to_date <- function(x) {
  sub_string <- str_sub(x, 1, 10)
  d <- mdy(sub_string)
  return(as.numeric(d))
}
floor_decade <- function(x){return(lubridate::year(x) - lubridate::year(x) %% 10)}

```


## Data cleaning
```{r}
ufo_sightings <- ufo_sightings %>%
  mutate(date = as_date(purrr::map_dbl(date_time, ~convert_to_date(.)))) # Convert to 'Date' format. Run only once, its slow af


# Some globals for ease
nightsky_img <- "data/nightsky2.jpg"
#font_files() %>% tibble() %>% filter(str_detect(family, "Showcard Gothic"))
font_add(family = "Showcard Gothic", regular = "SHOWG.TTF")
showtext_auto()

ufo <- ufo_sightings %>%
  filter(country == "us") %>% # Leaving only sightings in US
  filter(!(state %in% c("ak", "pr", "hi"))) %>% # Only mainland US
  select(date, code = state, description, encounter_length, latitude, longitude) %>%
  left_join(state_codes, by = "code") %>%
  mutate(decade = as.factor(purrr::map_dbl(date, ~floor_decade(.)))) %>% # Create decade variable
  drop_na(decade)


##### Cases by state in 2000-2010 ----
## Data preprocessing
by_state <- ufo %>%
    group_by(state, decade, .drop = F) %>%
    summarise(cases = n(),
              .groups = "drop")
  
by_state2 <- left_join(usa, by_state, by = c("region" = "state"), multiple = "all") %>%
  filter(decade %in% c(2000, 2010)) %>%
    left_join(uspop, by = "region")
  
cases_per_capita <- by_state2 %>%
  group_by(region) %>%
  summarise(cases = sum(cases), .groups = "drop") %>%
  left_join(uspop, by = "region") %>%
  mutate(cases_per_capita = cases/mean_pop)
  

by_state2 <- left_join(by_state2, select(cases_per_capita, region, cases_per_capita), by = "region")

```




# The Basics {.chapter}

## Background

This data includes over 80,000 recorded UFO "sightings" around the world, including the UFO shape, lat/long and state/country of where the sighting occurred, duration of the "event" and the `data_time` when it occurred.

Data comes originally from NUFORC, was cleaned and uploaded to Github by Sigmond Axel, and some exploratory plots were created by Jonathan Bouchet a few years back.


## Tables
Top 10 continental US states in reports oer capita:
```{r}
names(cases_per_capita) <- c("State", "Reports", "Mean population", "Reports per Capita")
cases_per_capita2 <- cases_per_capita %>%
  arrange(-`Reports per Capita`) %>%
  mutate(State = str_to_title(State))

flextable(head(cases_per_capita2, 10)) |>
  theme_vanilla() |>
  add_footer_lines("* Mean population = mean population size during the years 2000-2014") |>
  color(part = "footer", color = "#666666") |>
  align(part = "all", align = "center") |>
  autofit()
```

Bottom 10 continental US states in reports oer capita:
```{r}
names(cases_per_capita) <- c("State", "Reports", "Mean population", "Reports per Capita")
cases_per_capita2 <- cases_per_capita %>%
  arrange(`Reports per Capita`) %>%
  mutate(State = str_to_title(State))

flextable(head(cases_per_capita2, 10)) |>
  theme_vanilla() |>
  add_footer_lines("* Mean population = mean population size during the years 2000-2014") |>
  color(part = "footer", color = "#666666") |>
  align(part = "all", align = "center") |>
  autofit()
```

```{r, echo==FALSE}
glue::glue("Top state is {cases_per_capita2$State[1]}, with {round(cases_per_capita2[1,4], 3)} Reports per capita.
Bottom state is {cases_per_capita2$State[49]}, with {round(cases_per_capita2[49,4], 3)} Reports per capita.

Interestingly, Nevada - the home of area 51 - has only {round(cases_per_capita2$'Reports per Capita'[cases_per_capita2$State == 'Nevada'], 3)} reports per capita during this period.")

```


# Plotting {.chapter}
```{r}
plot <- ggplot(by_state2, aes(x = long, y = lat, fill = cases_per_capita, group = group)) +
  geom_polygon(color = "#00670c", show.legend = T) +
  scale_fill_gradient(low = "black", high = "#5dff00", limits = c(0, 0.2), breaks = seq(0, 0.2, length.out = 6), guide = guide_colorbar("Number of reported cases per capita", 
                                                                               title.position = "top",
                                                                               title.theme = element_text(color = "#5dff00", family = "serif"),
                                                                               title.hjust = 0.5,
                                                                               barwidth = 30,
                                                                               ticks.colour = NA)) +
  labs(title = "15 years of UFO sightings in the US between 2000 and 2014",
       caption = "Tomer Zipori | #TidyTuesday | Source: National UFO Reporting Center") +
  coord_fixed(1.3, clip = "off") +
  theme_minimal() +
  theme(plot.title = element_text(size = 24, vjust = -4, hjust = 0.5, color = "#5dff00", family = "Showcard Gothic"),
        plot.caption = element_text(color = "#5dff00", hjust = 1.05, family = "serif", size = 9),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom", legend.box = "horizontal", legend.text = element_text(color = "#5dff00", family = "mono", size = 14))
plot <- ggbackground(plot, nightsky_img)

plot

```

