---
title: "Advance Rmd"
output:
  html_document:
    toc: true
    # toc_float: true
# always_allow_html: true
---

```{r setup}
#| include = FALSE
knitr::opts_chunk$set(echo = FALSE)

library(afex)
library(glue)
library(scales)
```

## Equations / Math

We can write equations and math notation using LaTeX inside a pair of \$ symbols.

Here is an example of *displayed equations* (standalone):

$$
\hat{\sigma}^2_{x}=\frac{\sum_{i=1}^{N} (x_i-\bar{x})^2 }{N-1}
$$

We can also have *inline equations*, $t^2(df)=F(1,df)$ and $z^2=\chi^2(1)$.

## Glue

The `glue` package provides an easy interface for interpolate a strings. 

For example:

```{r glue}
#| echo=TRUE
glue::glue("{subject} {verb} {object}",
           subject = c("I","You"),
           verb = "love",
           object = "sleeping")

glue::glue("The mean of (0, 1, 2) is {mean(0:2)}")
```

Mixed with rmd, glue can be a powerful tool.

Let's look at how we may report an anova result:

```{r aov}
library(afex)

data(md_12.1)
a <- aov_ez("id", "rt", md_12.1,
            within = c("angle", "noise"), 
            anova_table = list(es = "pes"))
a
```

We want some inline reporting:

```{r format_with_glue}
library(scales) # function for formatting numbers and more

number2 <- number_format(accuracy = .01) # this makes a function that formats number to the 2 nearest decimals.
pvalue3 <- function(x) {
  p <- pvalue(x, accuracy = .001, add_p = TRUE)
  stringr::str_remove(p, "0(?=\\.)") # remove 0 before decimal 
}

 
aov_report <- 
  glue::glue_data(a$anova_table, 
                  "$F({number2(`num Df`)},{number2(`den Df`)})={number2(`F`)}$, ",
                  "$MSE={number2(MSE)}$, ",
                  "${pvalue3(`Pr(>F)`)}$, ",
                  "$\\eta^2_p={number2(pes)}$")
```

The main effect for *angle*: `r aov_report[1]`.  
The main effect for *noise*: `r aov_report[2]`.  
The interaction effect between *angle* and *noise*: `r aov_report[3]`.

```{r some_fig}
#| fig.width = 6,
#| fig.height = 2,
#| fig.cap = "Just a plot"
afex_plot(a, ~ noise, ~ angle, error = "within")
```

## Tables

Rmd comes with a basic function for formatting tables:

```{r make_table}
my_tab <- mtcars[1:5, 1:5]
knitr::kable(my_tab)
```

There are many other packages for advanced formatting, the most popular ones are:

- [`flextable`](https://davidgohel.github.io/flextable/)
- [`kableExtra`](https://haozhu233.github.io/kableExtra/) (partial support for Word outputs)
- [`gt`](https://gt.rstudio.com/) (no support for Word outputs)

*And [more](https://bookdown.org/yihui/rmarkdown-cookbook/table-other.html)...*

### flextable

```{r flextable}
library(flextable)

flextable(my_tab) |> 
  theme_alafoli() |> 
  color(~ drat > 3.5, ~ drat, color = "red") |> 
  add_header_row(
    values = c("some measures", "other measures"),
    colwidths = c(3, 2))
```

(Personally, this is my favorite package.)


### kableExtra

```{r kableExtra}
library(kableExtra)

kbl(my_tab) |> 
  kable_classic(full_width = F, html_font = "Cambria") |> 
  column_spec(6, color = "white",
              background = spec_color(mtcars$drat[1:8], end = 0.7),
              popover = paste("am:", mtcars$am[1:8]))

```

### gt

```{r gt}
library(gt)
gt(my_tab) |> 
  tab_header(
    title = md("Data listing from **gtcars**"),
    subtitle = md("`gtcars` is an R dataset")
  ) |> 
  tab_row_group(
    label = "Other cars",
    rows = 4:5
  ) |> 
   tab_row_group(
    label = "First 3 cars",
    rows = 1:3
  )
```


## Templates

...


## Exercise

1. Install the following packages 

```{r, eval=FALSE}
install.packages(c("rticles", "pagedown", "rmdformats"))
```

2. Make a new Rmd doc from a template.

Explore the template options. For example, if you get the following YAML:


```
---
output: pagedown::business_card
---
```

You can see all the options with `?pagedown::business_card`.

3. Load some dataset (come on, you have one).

4. Fill the doc with some data exploration
    - Make a plot
    - Make a table
    - Use `glue` to generate some text and/or equations.

